---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# arima2

<!-- badges: start -->
<!-- badges: end -->

The goal of `arima2` is to provide a set of tools to aid in the analysis of time series data in `R`. 
One such function within the `arima2` package is called `arima2`, which provides an interface to estimating Auto Regressive Integrated Moving Average (ARIMA) models using a random-restart algorithm. 
This function improves on the functionality of the `stats::arima` method, as it has the potential to increase the likelihood of the final output model. By design, the function cannot result in models with lower likelihoods than that of the `stats::arima` function. 
The potential for increased model likelihoods is obtained at the cost of computational efficiency, as the function is order $O(n)$ times slower than the `stats::arima` function, where $n$ is the number of random restarts.
Because the estimation of ARIMA models takes only a fraction of a second, we are of the opinion that potential to increase model likelihoods is well worth this computational cost.
The `arima2` function is implemented by modifying the source code of the `stats::arima` function. 

## TODO:

- [x] Implement `arima2` 
- [ ] Create simulate function for ARIMA models.
- [ ] Add datasets to the package 
- [ ] `ggplot2` figures for ARIMA objects - Dhajanae
- [ ] Create function that makes AIC table. 
- [ ] `probe`: a function to compare simulated data to a model 
- [ ] `auto.arima2:` implements that `auto.arima` function using `arima2`.
- [ ] (Jesse) `polyroots:` A function to get the roots of the polynomials created by the `AR` and `MA` coefficients of a model. 

## Installation

You can install the development version of arima2 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jeswheel/arima2")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(arima2)
## basic example code
```


You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>.
