---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# arima2

<!-- badges: start -->
<!-- badges: end -->

The goal of `arima2` is to provide a set of tools to aid in the analysis of time series data in `R`. 
One such function is `arima2::arima2`, which provides an interface to estimating Auto Regressive Integrated Moving Average (ARIMA) models using a random-restart algorithm. 
This function improves on the functionality of the `stats::arima` method, as it has the potential to increase the likelihood of the final output model.
By design, the function cannot result in models with lower likelihoods than that of the `stats::arima` function. 
The potential for increased model likelihoods is obtained at the cost of computational efficiency, as the function is order $O(n)$ times slower than the `stats::arima` function, where $n$ is the number of random restarts.
Because the estimation of ARIMA models takes only a fraction of a second, we are of the opinion that potential to increase model likelihoods is well worth this computational cost.
The `arima2` function is implemented by modifying the source code of the `stats::arima` function. 

## Installation

You can install the development version of arima2 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jeswheel/arima2")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(arima2)

# Set the seed for reproducible results. The seed was chosen so as to 
# demonstrate a reproducible example of arima2 outperforming stats::arima
# when the data are simulated from a known ARMA model. 
set.seed(48312)  

# Get model coefficients from ARMA(2, 2)
coefs <- sample_ARMA_coef(order = c(2, 2))

# Get model intercept 
intercept <- rnorm(1, sd = 50)

# Generate data from ARMA model 
x <- intercept + arima.sim(
  n = 100, 
  model = list(ar = coefs[grepl("^ar[[:digit:]]+", names(coefs))], 
               ma = coefs[grepl("^ma[[:digit:]]+", names(coefs))])
  )

# Plot the data
plot(x)

# Fit ARMA model using arima2 and stats::arima 
arma2 <- arima2(x, order = c(2, 0, 2))
arma <- arima(x, order = c(2, 0, 2))
```

In the example above, the resulting log-likelihood of the `stats::arima` function is `r round(arma$loglik, 2)`, and the log-likelihood of the `arima2` function is `r round(arma2$loglik, 2)`.
For this particular model and dataset, the random restart algorithm implemented in `arima2` improved the model likelihood by `r round(arma2$loglik - arma$loglik, 2)` log-likelihood units. 

## TODO:

- [x] Implement `arima2` 
- [ ] Create simulate function for ARIMA models.
- [ ] (Dhajanae) `ggplot2` figures for ARIMA objects
- [x] (Noel) Create function that makes AIC table. 
- [x] (Noel)`auto.arima2:` implements that `auto.arima` function using `arima2`.
- [x] (Jesse) `polyroots:` A function to get the roots of the polynomials created by the `AR` and `MA` coefficients of a model. 
